---
output: github_document
---


<!-- README.md is generated from README.Rmd. Please edit that file -->


<!-- badges: start -->
[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing) 
[![Travis build status](https://travis-ci.org/hypertidy/reproj.svg?branch=master)](https://travis-ci.org/hypertidy/reproj)
[![R build status](https://github.com/hypertidy/reproj/workflows/R-CMD-check/badge.svg)](https://github.com/hypertidy/reproj/actions)
[![CRAN status](https://www.r-pkg.org/badges/version/reproj)](https://cran.r-project.org/package=reproj)
[![CRAN\_Download\_Badge](http://cranlogs.r-pkg.org/badges/reproj)](https://cran.r-project.org/package=reproj)
[![R build status](https://github.com/hypertidy/reproj/workflows/test-coverage/badge.svg)](https://github.com/hypertidy/reproj/actions)
[![R build status](https://github.com/hypertidy/reproj/workflows/pkgdown/badge.svg)](https://github.com/hypertidy/reproj/actions)
<!-- badges: end -->
  


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# reproj

The goal of reproj is to reproject data between coordinate systems. 

The basic function `reproj()` takes input in diverse forms (matrix, data frame) and returns a 3-column matrix (or 4-column if `four = TRUE`), by transforming from map projection specfied by the  `source` argument to that specified by the `target` argument.

Matrix and data frame is not exactly massive diversity, but this model also allows adding methods for specific data formats that already carry a suitable `source` projection string. Currently we support types from the silicate and quadmesh and rgl packages, and only the `target` string need be specified. This model has obvious flexibility, for packages to import the generic and call it with the correct `source` (from the data format) and the `target` from user, or process controlled mechanism. 

The `source` argument must be named, and if it is not present a light check is made that the source data could be "longitude/latitude" and transformation to `target` is applied (this can be controlled by setting options). 

Until recently the `proj4` package was the only one available for generic data that will transform between arbitrary coordinate systems specified by _source_ and _target_ coordinate systems and with control over 'xy' versus 'xyz' input and output.  This package adds some further features by wrapping the need to convert longitude/latitude data to or from radians. 

Other R packages for transforming coordinates are not geared toward data that's not in a particular format. It's true that only GDAL provides the full gamut of available geographic map projections, but this leaves a huge variety of workflows and applications that don't need that level of functionality. 

## Dependencies

* The [PROJ](https://CRAN.r-project.org/package=PROJ) package is used preferentially if is functional, using the underlying 'PROJ-lib' at version 6 or higher. 
* The [proj4](https://CRAN.r-project.org/package=proj4) package is used if PROJ is not functional.

The proj4 package works perfectly well with the PROJ-lib at versions 4, 5, 6, or 7 and if this is preferred reproj can be set to ignore the PROJ R package (see [reproj-package](https://hypertidy.github.io/reproj/reference/reproj-package.html)). 



## Warning

Whether a particular source->target transformation pathway is appropriate for your application is your responsibility. 

## Installation

You can install the dev version of reproj from [github](https://github.com/hypertidy/reproj/) with:

```r
remotes::install_github("hypertidy/reproj")
```

## Example

This example shows how to convert between coordinate systems: 

```{r example}
library(reproj)
(pt <- (reproj(cbind(c(147, 148), c(-42, -45)), target = "+proj=laea +datum=WGS84", source = 4326)))

## to another coordinate system
(pt1 <- reproj(pt, target = "+proj=lcc +lat_1=-20 +lat_2=-10 +datum=WGS84", source = "+proj=laea +datum=WGS84"))

## and back again

reproj(pt1, target = "+proj=longlat +datum=WGS84", source = "+proj=lcc +lat_1=-20 +lat_2=-10 +datum=WGS84")

```

Note that the output is always in 'xyz' form, even if the z value is implied zero. The input may also be 'xyz' and it is the user's job to ensure this is specified correctly. It's not possible to automate every case, because among the three main cases: 

* longitude latitude and elevation,
* projected xy and elevation,
* geocentric xyz with embedded elevation (an offset to the radius `+a/+b` in PROJ terms),

it is not possible to easily capture the intended task. reproj sits at the simplest level of flexibility and control, and doesn't impose any restrictions of converting between these cases. 



## Contributions

Are you interested in reproj or even contributing? Great!  Please use the [issues](https://github.com/hypertidy/reproj/issues) tab for discussions. 


This is a wishlist / todo. 

* wrap modern PROJ in an R package with no other dependencies
* isolate the sf/lwgeom code for PROJ.4 in its own package
* allow choice of engine for `reproj.method` (this is already possible for your own classes)



Please note that the 'reproj' project is released with a [Contributor Code of Conduct](https://hypertidy.github.io/reproj/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.

